Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-07-04 article/article.tex
Commit 2024-07-04 22:14

The clear-sky reference values discussed above are based on the
climatological AOD and WC; hence they cannot describe accurately the
long-term variation of \(\text{GHI}_\text{ref}\) due to long-term
changes in the two atmospheric constituents, mainly AOD. As reported by
\citet{Natsis2023}, there is a long-term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky
data was attributed to long-term changes in aerosol effects. Therefore,
an adjustment of the \(\text{GHI}_\text{ref}\) during the period of
study was made using simulation with the RTM based on the long-term
variations of the AOD. As AERONET data start only in 2003, we used for
the period 1993 -- 2005 estimates of changes in AOD at
\(340\,\text{nm}\) derived from a collocated Brewer spectrophotometer
\citep{Kazadzis2007} to calculate the trend in \(\text{GHI}_\text{ref}\)
during this period.

According to this study, in the period 1997 -- 2005 the mean AOD at
\(340\,\text{nm}\) is \(0.403\) with a trend of \(-3.8\pm0.93\,\%\) per
year, corresponding to a change of \(0.0153\) per year. Using an
Ångström coefficient \(\alpha = 1.6\), this translates to a change in
the Ångström coefficient \(\beta=0.00272\) per year (or \(\beta=0.084\)
in 1997 and \(\beta=0.059\) in 2005). Simulations with uvspec for the
above Ångström coefficients, with WC of \(15.7\,\text{mm}\) and
\(15.4\,\text{mm}\) for 1997 and 2005 respectively, also taken from
AERONET, and for a SZA of \(55^\circ\) reveal a trend of \(+0.21\,\%\)
per year in \(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was
chosen as representative of all days in the year in order to get a rough
estimate of the annually averaged change in clear sky irradiance. For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC
from AERONET in a similar simulation scheme to calculate the monthly
mean of clear sky irradiance, and finally the trend of \(+0.14\,\%\) per
year. We applied these two long-term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

\caption{Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD.}\label{fig:CS-change}
\end{figure}

@@

In this study our main focus was to quantify over irradiance events
related to CEs. A key issue for achieving this goal is to define a
threshold for the CE identification, representative of the clear-sky
irradiance at the time of each GHI measurement. This depends on the
selection of the appropriate atmospheric parameterization for the RTM
simulations. The implementation of the long-term change of AOD,
discussed in section\nobreakspace{}\ref{rtmcs}, allows capturing a large
part of the natural variability of clear-sky GHI. However, the
short-term variability of AOD cannot be taken adequately into account
when using monthly values in the model simulations. We tried different
approaches in order to strengthen the robustness of the methodology and
to compensate for the limited accuracy of the RTM input data and the
unpredictable natural variability of the atmosphere.

>>>>

The clear-sky reference values discussed above are based on the
climatological AOD and WC; hence they cannot describe accurately the
long-term variation of \(\text{GHI}_\text{ref}\) due to long-term
changes in the two atmospheric constituents, mainly AOD. As reported by
\citet{Natsis2023}, there is a long-term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky
data was attributed to long-term changes in aerosol effects. Therefore,
an adjustment of the \(\text{GHI}_\text{ref}\) during the period of
study was made using simulations with the RTM based on the long-term
variations of the AOD. As AERONET data start only in 2003, we used for
the period 1993 -- 2005 estimates of changes in AOD at
\(340\,\text{nm}\) derived from a collocated Brewer spectrophotometer
\citep{Kazadzis2007} to calculate the trend in \(\text{GHI}_\text{ref}\)
due to aerosols during this period.

According to \citet{Kazadzis2007}, in the period 1997 -- 2005 the mean
AOD at \(340\,\text{nm}\) is \(0.403\) with a trend of
\(-3.8\pm0.93\,\%\) per year, corresponding to a change of \(0.0153\)
per year. Using an Ångström coefficient \(\alpha = 1.6\), this
translates to a change in the Ångström coefficient \(\beta\) of
\(0.00272\) per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in
2005). Simulations with uvspec for the above Ångström coefficients, with
WC of \(15.7\,\text{mm}\) and \(15.4\,\text{mm}\) for 1997 and 2005
respectively, also taken from AERONET, and for a SZA of \(55^\circ\)
reveal a trend of \(+0.21\,\%\) per year in \(\text{GHI}_\text{ref}\).
The SZA of \(55^\circ\) was chosen as representative of all days in the
year in order to get a rough estimate of the annually averaged change in
clear sky irradiance. For the period 2005 -- 2023 we used the mean
monthly values of AOD and WC from AERONET in a similar simulation scheme
to calculate the monthly mean clear-sky irradiance, and finally the
trend of \(+0.14\,\%\) per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

\caption{Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1992 -- 2023.}\label{fig:CS-change}
\end{figure}

@@

In this study our main focus was to quantify the over irradiance related
to CEs. A key issue for achieving this goal is to define a threshold for
the CE identification, representative of the clear-sky irradiance at the
time of each GHI measurement. This depends on the selection of the
appropriate atmospheric parameterization for the RTM simulations. The
implementation of the long-term change in AOD, discussed in
section\nobreakspace{}\ref{rtmcs}, allows capturing a large part of the
natural variability of clear-sky GHI. However, the short-term
variability of AOD cannot be taken adequately into account when using
monthly values in the model simulations. We tried different approaches
in order to strengthen the robustness of the methodology and to
compensate for the limited accuracy of the RTM input data and the
unpredictable natural variability of the atmosphere.
